\documentclass[12pt,letterpaper]{article}
\usepackage{geometry}
\usepackage{enumerate}
\usepackage{amsmath,amsthm,amssymb,amsfonts}

\title{AE504: Homework Assignment 1}
\author{Alex Faustino}


\begin{document}
\maketitle
\thispagestyle{empty}

\begin{enumerate}
    \item[\textbf{Problem 1}.] {
    A single-variable polynomial either achieves a minimal value, or its infimum equals $-\infty$. However, a two-variable polynomial can be bounded from below while still not ever achieving its infimum. Find an example of such a polynomial. \\
    (Hint: construct a polynomial that is always strictly positive, but can attain values arbitrarily close to $0$.)
    }
    \item[\textbf{Solution:}] {
    Consider a polynomial function $f:\mathbb{R}^2\to\mathbb{R}$ that is given by \begin{equation*}
        f(x,y)=(1-xy)^2+x^2
    \end{equation*}
    Because of the squared terms, $f$ is clearly strictly positive. We're interested in the set of solutions, $A$, that satisfy $xy=1$ where $y \gg x$. As $x\to0$, $f$ approaches but never achieves $0$, so $0<a'$ for all $a'\in A$. And there exists an $a'$ for all $\epsilon > 0$ s.t. $a'< 0 + \epsilon$. Therefore, $f$ is bounded below by $\inf f = 0$ that it never achieves.
    }
    
    \item[\textbf{Problem 2}.] {
    Let $f:\mathbb{R}^2\to\mathbb{R}$ be given by $f(x,y)=2x^3+3y^2+3x^2y-24y$. 
    \begin{enumerate}
        \item[\textbf{(a)}] Find all local minima of $f$ or prove that no such minima exist.
        \item[\textbf{(b)}] Find all global minima of $f$ or prove that no such minima exist.
    \end{enumerate}
    }
    \item[\textbf{Solution:}] {
    We find local minima by first identifying candidate points where $\nabla f(x,y) = 0$. 
    \begin{equation*}
        \nabla f(x,y) = \begin{bmatrix} 6x^2+6xy \\ 6y^2+3x^2-24\end{bmatrix}
    \end{equation*}
    We find that $\nabla f(x,y) = 0$ at four points: (0,$\pm 2$), ($\frac{2\sqrt{6}}{3}$, $-\frac{2\sqrt{6}}{3}$), and ($-\frac{2\sqrt{6}}{3}$, $\frac{2\sqrt{6}}{3}$). If the Hessian of $f$, $Hf(x,y)$, is positive definite at a candidate point then it is a local minimum. We determine that $Hf(x,y)>0$ if all its eigenvalues have positive real part.
    \begin{equation*}
        Hf(x,y) = \begin{bmatrix} 12x+6y & 6x \\ 6x & 12y \end{bmatrix}
    \end{equation*}
    }
    We find that only $Hf(0,2)>0$ therefore $f(x,y)$ has one local minimum $f(0,2)=-36$. We also find that $Hf(0,-2)<0$ and that as $x, y \to -\infty$ so does $f(x,y)$ meaning a global minimum does not exist.

    \item[\textbf{Problem 3}.] {
    Let $f:\mathbb{R}^2\to\mathbb{R}$ be defined by $f(x,y)=\sin^2(x)+\arctan^3(xy)$. For every positive integer $n$ and every $M\geq 0$, find $$\min_{(x,y)\in S_{n,M}} f(x,y)\textrm{,}$$ where $S_{n,M}=[0,n \pi]\times [-M,M]$, and determine where this minimum is achieved.
    }
    \item[\textbf{Solution:}] {
    We begin our analysis by noting that when unconstrained $\sin^2(x)$ is bounded by $[0,1]$, has a minimum of $0$, and that minimum is achieved at $0$ and every integer multiple of $\pi$. We also note that when unconstrained $\arctan^3(xy)$ is bounded by $(-\frac{\pi^3}{8},\frac{\pi^3}{8})$ and has no minimum but an infimum of $-\frac{\pi^3}{8}$ which is never achieved.
    
    It is then clear that when $f(x,y)$ is constrained to $S_{n,M}=[0,n \pi]\times [-M,M]$ that it will achieve its minimum when $\sin^2(x)=0$ and $\arctan^3(xy)\to -\frac{\pi^3}{8}$. Therefore, $\min f(x,y)=\arctan^3(-n \pi M)$ for all positive integer $n$ and every $M\geq 0$.
    }
    
    \item[\textbf{Problem 4}.] {
    As discussed in class, the global maximum of $f(x,y)=xy(1-x-y)$ on $S=\{(x,y,z)~|~x\geq 0, y\geq 0, x+y\leq 1\}$ can be found applying the first-order condition and finding the maximum among the four candidates. However, it can be shown that $\nabla^2 f(x,y)\not\geq 0$ for all candidate points. Nevertheless, $f$ has a global minimum on $S$. 
    \begin{enumerate}
        \item[\textbf{(a)}] Explain why the same method that worked for a maximum does not work for a minimum.
        \item[\textbf{(b)}] Find $\min_{(x,y)\in S}f(x,y)$ and determine all points at which it is achieved.
    \end{enumerate}
    }
    \item[\textbf{Solution:}] {
    The second-order condition was satisfied for the global maximum on $S$ because it also a local maximum for the unconstrained problem. This is not the case for the global minimums on $S$. As we discussed in class, $\min_{(x,y)\in S}f(x,y)=0$ on the lines $x=0$, $y=0$, and $x+y=1$. Since minimums on $S$ are achieved on the bounds of the set it is only necessary that their directional derivatives are increasing towards the interior.
    }
\end{enumerate}

\end{document}